---
title: "DS02_LAB5"
author: "Caroly Coronado- Vargas"
date: "11/17/2021"
output: html_document
---

```{r}
library(ggplot2)
library(tidyverse)
library(lubridate)
library(maps)

acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)

pers<- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)

pers
acc
```

## 1 ##

```{r}
df1<- pers%>%
  filter(INJ_SEV==4)

df1
```

## 2 ## 

```{r}
## tells us more dangerous/fatal cars in each state

DANGERVEH<-df1%>%
  filter(!is.na(MAKE))%>%
  group_by(MAKE, STATE)%>%
  summarize(FatalCount= n())%>%
  arrange(desc(FatalCount))%>%
  group_by(STATE)%>%
  filter(FatalCount ==max(FatalCount))
  
  
DANGERVEH

```
## 3 ##

*The most dangerous vehicle was defined as a vehicle with the most counts (n) of level 4 Injury severity (INJ_SEV), omitting those which "NA" as car model entries*

```{r}
states<- readxl::read_xlsx("FRPP_GLC_-_United_StatesNov42021.xlsx")
states

statesmap <- map_data("state")%>%
  mutate(
    region = toupper(region)
    )%>%
  left_join(states %>%
              mutate(
                 `State Code` = as.integer(`State Code`)
                 )%>% select(`State Name`, `State Code`)%>% distinct, by = 
c(region = 'State Name'))%>%
  group_by(`State Code`)%>%
  summarize(
    long = mean(long), 
    lat = mean(lat)
    )%>%
  left_join(DANGERVEH, by = c('State Code' = 'STATE'))
  
map_data("state")%>%
  ggplot(aes(x = long, y = lat)) + geom_path(aes(group=group)) + geom_text(aes(label=MAKE), data=statesmap)

```
## 4 ##

```{r}
joined<- left_join(pers, acc, by= "ST_CASE")

joined
```
## 5 ## 

*Most accident happen Friday-Sunday with a peak on Saturday. The peak time for accident is during the night. Men are almost double the accident reported.*

```{r}
#day
joined%>%
  group_by(DAY_WEEK)%>%
  summarize(TotalAcc=n())%>%
  ggplot(aes(x=DAY_WEEK, weight = TotalAcc))+geom_bar()

#hour
joined%>%
  group_by(HOUR.y)%>%
  summarize(TotalAcc=n())%>%
  ggplot(aes(x=HOUR.y, weight = TotalAcc))+geom_bar()

#sex
joined%>%
  group_by(SEX)%>%
  summarize(TotalAcc=n())%>%
  filter(!SEX>2)%>%
  ggplot(aes(x=SEX, weight = TotalAcc))+geom_bar()
```
## 6 ## 

*We can see that most accidents happen in bigger cities specially alot of California* 

```{r}
df3<-df1%>%
  group_by(STATE, COUNTY)%>%
  count()

county <- map_data("county")%>%
  mutate(
    region = toupper(region), subregion = toupper(subregion)
    )%>%
  left_join(states %>%
              mutate(
                 `State Code` = as.integer(`State Code`),`County Code` = as.integer(`County Code`)
                 )%>% select(`State Name`, `State Code`, `County Name`, `County Code`)%>% distinct, by = 
c(region = 'State Name', subregion= 'County Name'))%>%
  left_join(df3, by = c('State Code' = 'STATE', 'County Code' = 'COUNTY'))
  

county%>%
  ggplot(aes(long, lat)) + geom_polygon(aes(group= group, fill = n))
```



